---
title: "Statistical Inference - Inferential Data Analysis"
author: "Silva, RAFAEL"
date: "December 4, 2017"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This section presents all the libraries used during the project.

```{r, message=FALSE}
library(datasets)
library(ggplot2)
library(dplyr)
library(knitr)
```

## Introduction

The objective of this project is to make some inferences about the length variation on each supplement and dose on the ToothGrowth data from the R datasets package. The análysis is going to follow the four steps below:

1. Exploratory data analysis and summary;
1. Hypothesis testing;
1. Conclusions and assumptions.

The chunk bellow is used to read the data set into the workspace.

```{r}
data("ToothGrowth")
```

For more information on the data set (description, dimensions, variables and source) enter ?ToothGrowth on the console to check its documentation.

## Exploratory analysis

To start the analysis, a brief verification on how the dataset looks like is necessary.

```{r}
dim(ToothGrowth)
```

```{r}
head(ToothGrowth)
```

As the outputs above show, the data set has 60 observations on 3 variables. As the instroduction chapter states, the objective is to make some inferences about the length variation on each supplement and dose. The boxplot below gives a feeling about how each combination affects the length.

```{r}
ToothGrowth <- ToothGrowth %>%
      mutate(suppDose = paste(supp, format(dose, digits = 2), sep = "/"))

g1 <- ggplot(ToothGrowth, aes(x = suppDose, y = len)) + 
      geom_boxplot() + 
      labs(title = "Length by Supplement and Dose", 
           x = "Supplement/Dose", y = "Length")

print(g1)
```

It is important to notice that, in the graph above, the x axis does not present a continuous scale. This means that when analysing the graph you should not look at how the length increase "evolve" through the doses but at how each group behave in the y scale.

In the table below, the main statistical attributes of each group are summarized.

```{r, fig.width=5, fig.height=3}
summ <- ToothGrowth %>%
      group_by(suppDose) %>%
      summarise(min = min(len),
                q1 = quantile(len, .25),
                mean = mean(len), 
                median = median(len),
                q3 = quantile(len, .75),
                max = max(len),
                variance = var(len), 
                n = length(len))

kable(summ, caption = "Data Summary by Supplement and Dose", 
      col.names = c("Supp/Dose", "Min", "1st Qu.", "Mean", "Median", "3rd Qu.",
                    "Max", "Variance", "n"))
```

Continue...